---
title: "esm206_assignment4"
author: "Claire Meuter"
date: "11/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(here)
library(effsize)
library(broom)
library(ggbeeswarm)
library(kableExtra)
library(lubridate)
```

### Introduction 
A 4 - 5 sentence "mini-introduction" to what this study is exploring, including basic background information about the subject, site and study (with citations as needed), and a brief overview of what is included in the exploratory report. You should probably do this last so you know what’s contained in the report.


## Data and analyses 
A 3 - 4 sentence section in which you briefly describe the data, including citation & credit as needed, and provide an overview of the main analysis, statistical methods and tools used.

## Juvenile hare initial investigation: 
# Annual juvenile hare trap counts
Count the total number of juvenile hare trappings during each year of the study, and create a finalized data visualization of the counts by year. Include a figure caption below the figure in your report. 

i.)
# converted date to date class and created year column
```{r}
bonanza_hares <- read.csv(here("data", "bonanza_hares.csv"))
juvenile_inv <- bonanza_hares %>%
  filter(age == "j") %>%  #filtered to only keep juvenile samples in new data set
  mutate(date_new = mdy(date)) %>% # converted dates to date class
  mutate(year = year(date_new)) %>% # added new column
  mutate(sex = case_when( # edited sex to show on data as male and female
    sex == "f" ~ "female", 
    sex == "m" ~ "male")) %>%
      mutate(grid = case_when( # edited site names to show as full names 
        grid == "bonbs" ~ "Bonanza Black Spruce",
        grid == "bonmat" ~ "Bonanza Mature",
        grid == "bonrip" ~ "Bonanza Riparian"
      ))
```
 # counted my data by year
```{r}
juvenile_counts <- juvenile_inv %>%
  group_by(year) %>%
  summarise(counts = n())
```

```{r}
ggplot(data = juvenile_inv, aes(x = year)) + 
  geom_histogram(bins = 15, color = "black", fill = "darkolivegreen") + scale_x_continuous(breaks = c(1998, 2000, 2002, 2004, 2006, 2008, 2010, 2012)) + labs(title = "Annual Counts of Trapped Juveniles", x = "Year", y = "Count")
```

Figure 1: 
Total number of juveniles trapped from 1998 to 2012. 

What were the minimum and maximum counts of juvenile hare trappings?
We see the maximum number of juveniles 
Mean or median annual number of juvenile hares trapped?
General/notable trends?
These are absolute counts, but this count is also impacted by effort (e.g. the number of days and number of traps in the study each year). What would you suggest be done moving forward (you are not expected to actually do this) to standardize the juvenile hare population in future analyses?

### ii.) Visualize juvenile hare weights.
juvenile hare weights by sex and site (grid)


```{r}
ggplot(data = juvenile_inv) + geom_jitter(aes(x = sex, y = weight, color = sex)) + 
  facet_wrap(~ grid)
```



### iii.) Juvenile weight comparison (male & female snowshoe hares). 
Further compare mean weights for **juvenile male and female snowshoe hares** included in this study. In this section, you should include:

A finalized table containing the mean, standard deviation, and sample size for male and female juvenile snowshoe hares, with a table caption (remember, a table caption should be placed *above* the table in your report)
A 2 - 3 sentence comprehensive statement of the means comparison, including at least:
The actual difference in means
Effect size (Cohen's *d*) (see effsize::cohen.d() example in the Lab 5 key)
Outcome of an unpaired two-sample t-test --- lab 6 

```{r}
juvenile_table <- juvenile_inv %>% 
  group_by(sex) %>% 
  summarize(
    mean_weights = round(mean(weight, na.rm = TRUE),2),
    sd_weight = round(sd(weight, na.rm = TRUE),2),
    n = n()
  )
juvenile_table %>%
  kable(col.names = c("Sex", 
                      "Mean Weights", 
                      "Standard Deviation", 
                      "Sample Size"), 
        caption = "Juvenile Weights by Sex") %>%
  kable_styling(full_width = FALSE) #kable_styling(full_width = FALSE) spaces columns out
```
A 2 - 3 sentence comprehensive statement of the means comparison, including at least:
The actual difference in means
Effect size (Cohen's *d*) (see effsize::cohen.d() example in the Lab 5 key)
Outcome of an unpaired two-sample t-test --- lab 6 

# Split off into groups, pull sample vectors
```{r}
weight_female <- juvenile_inv %>%
  filter(sex == "female") %>%
  drop_na(weight) %>%
  pull(weight)

weight_male <- juvenile_inv %>%
  filter(sex == "male") %>%
  drop_na(weight) %>%
  pull(weight)
```

```{r}
juvenile_t <- t.test(weight_male, weight_female)

juvenile_t
```
There is a negligible probability of randomly selecting two samples from populations with the same mean that are this different by chance. They are almost certainly from populations with different mean mpgs - which makes sense. Of course they do. More interesting: How different are they?

There is a low probability (less than 0.05%) of randomly selecting two samples from populations with the same mean that are this different by chance. While it is possible These are very likely from populations with different mean weights. Based on our data table above this makes sense. 

How to do actual difference of means?
```{r}
juvenile_weight_cohen <- cohen.d(weight_male, weight_female)

juvenile_weight_cohen # to view
```
The effect size was small (Cohen's *d* = `r juvenile_weight_cohen$estimate`).

Writing it all out: Mean weight for males juvenile snowshoe hares (`r round(juvenile_t$estimate[1],2)` grams) is greater than that for female juvenile (`r round(juvenile_t$estimate[2],2)` grams); the effect size is small (Cohen's *d* = `r round(juvenile_weight_cohen$estimate,2)`) and the difference significant (two-sample t-test, t = `r round(juvenile_t$statistic,2)`, p < 0.007)

# Visual data exploration with histograms and QQ-plots

```{r}
juvenile_male_female <- juvenile_inv %>% 
  filter(sex %in% c("female", "male"))
```



```{r, include=FALSE}
ggplot(data = juvenile_male_female, aes(x = weight)) +
  geom_histogram(bins = 15) +
  facet_wrap(~sex)

ggplot(data = juvenile_male_female, aes(sample = weight)) +
  geom_qq() +
  facet_wrap(~sex)
```






## iv.) Relationship between juvenile weight & hind foot length
Explore the relationship between juvenile snowshoe hare hind foot length (variable `hindft`, recorded in millimeters), and weight. Include all juvenile hares in the study. Create a final visualization of juvenile hare hind foot length versus weight. Include a figure caption.


```{r}
weight_foot_length <- ggplot(data = juvenile_inv) + 
  geom_point(aes(x = hindft, y = weight), color = "dodgerblue4") +
  labs(x = "Hind Foot Length (mm)",
       y = "Juvenile Hare Weights (g)",
       title = "Juvenile Hare Hind Foot Length by Juvenile Hare Weights" )

weight_foot_length
```


